Synthesizing Quotient Lenses

### 2018-03-10 papers/QuotientLenses/main.tex
finished pass over intro

Our solution to this shortcoming is to introduce {\em Quotient Regular
Expressions} (QREs), which enable programmers to simultaneously specify a
regular expression and a set of canonical representatives for matching
strings.  Given an QRE, we can automatically infer the corresponding
canonization functions.

@@

Our solution to this problem is to introduce {\em QRE lenses}, which
are a new class of quotient lens that can be automatically
synthesized, obviating the need to write lenses by hand.  A QRE lens
is like a normal lens except it maps between formats described by
QREs. We prove that every composition of QRE lenses can be rewritten
to be the composition of a source canonizer, a bijective lens, and a
target canonizer. This normalization property means we can synthesize
QRE lenses by extending an existing algorithm for synthesizing
bijective lenses~\cite{optician}.

Generating a QRE lens requires only a pair of QREs to describe the
source and target formats and a (possibly empty) suite of examples
demonstrating the mapping.  For example, the following code

\begin{lstlisting}
let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
\end{lstlisting}

binds \codefont{l} to a synthesized QRE lens mapping between names in
the comma-separated form describd by the QRE \codefont{comma_name}
and the space-separated form described by the
QRE \codefont{space_name}.

Given this information, we can exploit and
extend recent work on algorithms for synthesis of bijective
lenses~\cite{optician} to generate quotient lenses. The lenses generated by
our synthesis algorithm are guaranteed to translate data back and forth
between source and target QREs and to satisfy the provided examples.
Moreover, while our synthesized quotient lenses have a relatively
restricted form (which cuts down the amount of non-determinism in
a specification and greatly speeds up synthesis), we prove that
this rigidity does not compromise their expressiveness.

The following list summarizes the central contributions of our work.
\begin{enumerate}
  \item We introduce {\em Quotient Regular Expressions} (QREs),
  a compact, convenient notation for simultaneously defining a
  regular language modulo some equivalence relation and a canonizer
  for that relation (Section~\ref{QRE}).
  \item We introduce {\em QRE lenses} which are lenses that map data specified
  using QREs.
  In addition, we will descibe a particular set of QRE lens combinators and
  prove that QRE lenses defined using these combinators can be transformed to
  a QRE lens that us defined using just two of these combinators; this is the
  main technical contribution of this paper.
  (Section~\ref{QRE-lenses}).
  \item We use this result to reduce the problem of {\em synthesizing} QRE
  lenses to the problem of synthesizing bijective string lenses which we studied
  in~\cite{optician}(Section~\ref{synth}).
  \item We extend the Boomerang {\em implementation} with QREs
  and QRE lens synthesis.  We demonstrate the utility of our implementation by
  using it to synthesize QRE lenses between data formats drawn from the
  {\tt data.gov} database (Section~\ref{impl}).
\end{enumerate}
Sections~\ref{relwork} and~\ref{concl} discuss related and future work.

\bcp{We've said that one of the main contributions of the paper is
synthesizing lenses, but we don't actually show this happening in the
introduction.  Should we?}

>>>>

Our solution to this shortcoming is to introduce {\em Quotient Regular
Expressions} (QREs), which enable programmers to simultaneously specify a
regular expression and a set of canonical representatives for matching
strings.  Given an QRE, we can automatically infer the corresponding
canonization functions.  
\ksf{Define comma\_name here as a really simple example? }

@@

Our solution to this problem is to introduce {\em QRE lenses}, which
are a new class of quotient lens that can be automatically
synthesized, obviating the need to write lenses by hand.  A QRE lens
is like a normal lens except it maps between formats described by
QREs. We prove a normal form theorem (Theorem~\ref{normal form}) that says that every
composition of QRE lenses can be rewritten 
to be the composition of a source canonizer, a bijective lens, and a
target canonizer. This normalization property means we can synthesize
QRE lenses by extending an existing algorithm for synthesizing
bijective lenses~\cite{optician}.

Given this framework, generating a QRE lens requires only a pair of QREs to describe the
source and target formats and a (possibly empty) suite of examples
demonstrating the mapping.  For example, the following code

\begin{lstlisting}
let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
\end{lstlisting}
binds \codefont{l} to a synthesized QRE lens mapping between names in
the comma-separated form describd by the QRE \codefont{comma\_name}
and the space-separated form described by the
QRE \codefont{space\_name}. 
The lenses generated by
our synthesis algorithm are guaranteed to translate data back and forth
between source and target QREs and to satisfy the provided examples.
Our synthesized quotient lenses have a relatively
restricted form (which greatly speeds up synthesis), but the 
normalization theorem means 
this restriction does not compromise expressiveness.

The following list summarizes the contributions of our work.
\begin{enumerate}
  \item We introduce {\em Quotient Regular Expressions} (QREs), which are
  a compact, convenient notation for simultaneously defining a
  regular language modulo some equivalence relation and a canonizer
  for that relation (Section~\ref{QRE}).
  \item We introduce {\em QRE lenses}, which are lenses that translate
  between data formats specified using QREs. (Section~\ref{QRE-lenses}).
  \item We descibe a particular set of QRE lens combinators and prove
  that QRE lenses defined using these combinators can be normalized
  (Theorem~\ref{normal form}), which is the main technical
  contribution of this paper.
  (Section~\ref{QRE-lenses}).
  \item We use this result to reduce the problem of {\em synthesizing} QRE
  lenses to the problem of synthesizing bijective string lenses
  previously studied in~\cite{optician}(Section~\ref{synth}).
  \item We extend Boomerang with QREs and QRE lens synthesis. 
  We use the resulting implementation to demonstrate the utility of
  our approach by synthesizing QRE lenses between
  data formats drawn from the {\tt data.gov} database (Section~\ref{impl}).
\end{enumerate}
Sections~\ref{relwork} and~\ref{concl} discuss related and future work.
